SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-10-29 docs/questioning.tex
first draft of parse-tree validation, also added new subsection: Hybrid

Figure~\ref{fig:defenses} presents a taxonomy of the
defenses that protect web applications against attacks based on
code injection vulnerabilities.
We can identify two main approaches, namely:
{\it design-based} and {\it behavior-based}.
The former involves mechanisms that make applications
immune to different attack vectors by design~\cite{JL75,L81}.
The latter incorporates a variety of methods and
schemes that inspect the behavior of applications
and detect attacks based on specific criteria~\cite{D76,A00}.
Both categories involve more than one subcategories.
For every subcategory we provide elements regarding
every dimension described in Section~\ref{sec:dimensions}.

@@

The key idea behind this approach, is to compare
the tree representation of the abstract syntactic
structure of the executable source code that is
about to be executed with the one hat was originally intended.
If the trees diverge, the application is probably under attack.

In the case of {\sc dsl} injection attacks, mechanisms check 
the query before the inclusion of user input with the one
resulting after the inclusion of user input.

@@

{\it {\sc sql}rand}~\cite{BK04} is based on {\sc isr}
to detect {\sc sql} injections in the following manner:
initially, it allows programmers to create {\sc sql} statements
using randomized instructions instead of standard keywords.
The modified queries are either reconstructed at runtime using
the same key that is inaccessible to the malicious user,
or the user input is tagged with delimiters that allow
an augmented {\sc sql} grammar to detect the
attack. {\sc sql}rand is one of the few mechanisms that
detect {\sc sql} injection attacks on the database
level (locus $\Omega$).

@@

{\sc isr} is a deterministic approach that can be applied
to detect different attacks in an efficient manner.
However, Sovarel et al.~\cite{SEP05}
have investigated thoroughly the effectiveness
of {\sc isr} and showed that a malicious user
may be able to circumvent it by determining
the randomization key and their results
indicate that doing {\sc isr} in a way that
provides a certain degree of security against a motivated
attacker is more difficult than previously thought.

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.52]{defenses.pdf}
\end{center}
\caption{\label{fig:defenses}The basic categories of web application
attack countermeasures ({\sc draft}).}
\end{figure*}

@@

  &   {\it {\sc sql}rand}~\cite{BK04} & 286 & (3,{\sc na},{\sc na},0)\_a & +6.5{\it ms} & {\sc sql} injection \\ 

@@

>>>>

\begin{figure*} [ht]
\begin{center}
\leavevmode
\includegraphics[scale=0.62]{defenses.pdf}
\end{center}
\caption{\label{fig:defenses}The basic categories of web application
attack countermeasures ({\sc draft}).}
\end{figure*}

Figure~\ref{fig:defenses} presents a taxonomy of the
defenses that protect web applications against attacks based on
code injection vulnerabilities.
We can identify three main approaches, namely:
{\it design-based}, {\it behavior-based}
and {\it hybrid}.
The design-based category involves mechanisms that make applications
immune to different attack vectors by design~\cite{JL75,L81}.
The behavior-based incorporates a variety of methods and
schemes that inspect the behavior of applications
and detect attacks based on specific criteria~\cite{D76,A00}.
Both categories involve more than one subcategories.
The hybrid category borrows elements from more than one subcategories.
For every subcategory we provide elements regarding
every dimension described in Section~\ref{sec:dimensions}.

@@

The key idea behind this approach, is to compare
the tree representation of the abstract syntactic
structure of the executable source code
with the one that was originally intended.
If the trees diverge, the application is probably
under attack.

In the case of {\sc dsl} injection attacks, mechanisms check 
the query before the inclusion of user input with the one
resulting after the inclusion of user input.
Both mechanisms that implement this approach,
{\sc sqlg}uard~\cite{BWS05} and
{\sc sql}check~\cite{SW06} are pretty similar
and detect the attack before a query reaches the
database (locus {\sc z}).
Contrary to {\sc sqlg}uard~\cite{BWS05} (and many other
mechanisms), {\sc sql}check~\cite{SW06} has been
extensively tested in terms of diagnostic performance
as seen in Table~\ref{tab:comp}. A disadvantage of
those mechanisms is that the application must be modified
every time a query is about to be sent to a database
for execution.

As we will see in Subsection~\ref{sec:hybrid},
there are mechanisms that borrow elements from
this approach and examine the syntax trees
of scripts to detect JavaScript-driven {\sc xss}.

@@

{\it {\sc sql}rand}~\cite{BK04} is based on {\sc isr}
to detect {\sc sql} injections in the following manner:
initially, it allows programmers to create {\sc sql} statements
using randomized instructions instead of standard keywords.
The modified queries are reconstructed at runtime using
the same key that is inaccessible to the malicious user.
{\sc sql}rand is one of the few mechanisms that
detect {\sc sql} injection attacks on the database
level (locus $\Omega$).

@@

{\sc isr} is a deterministic approach that can be applied
to detect different attacks in an efficient manner.
However, Sovarel et al.~\cite{SEP05}
have investigated thoroughly the effectiveness
of {\sc isr} and showed that a malicious user
may be able to circumvent it by determining
the randomization key and their results
indicate that doing {\sc isr} in a way that
provides a certain degree of security against a motivated
attacker is more difficult than previously thought.
Furthermore, developers who wish to use
such mechanisms must follow good coding practices
and never show a randomized code statement in any case
(i.e. an exception error) because in this way
they will reveal the encoding key.

@@

\subsection{Hybrid}
\label{sec:hybrid}

{\sc xss-guard}~\cite{BV08}
maps legitimate scripts to {\sc http} responses.

@@

  &   {\it {\sc sql}rand}~\cite{BK04} & 286 & (3,{\sc na},{\sc na},0)\_a & $\le$6.5{\it ms} & {\sc sql} injection \\ 

@@

  \multirow{1}{*}{Hybrid}
